---
title: "P8105_hw1_tj2383"
author: Tanvi Jain
date: 2018-09-15
output: html_document
---

Loaded necessary packages:

```{r}
library(tidyverse)
```

Question 1:
I created a data frame (q1_df) comprised of:

-A random sample (random_samp) of size 10 from a uniform[0, 5] distribution
-A logical vector (logical_vec) indicating whether elements of the sample are greater than 2
-A (length-10) character vector (character_vec)
-A (length-10) factor vector (factor_vec)
```{r question_1_code}
set.seed(168)
q1_df = tibble(
  random_samp = runif(10, min = 0, max = 5),
  logical_vec = random_samp > 2,
  character_vec = c("dog", "cat", "rabbit", "frog", "mouse", "rat", "fly", "hippo", "horse", "bird"),
  factor_vec = factor(c("low", "medium", "high", "medium", "low", "medium", "high", "low", "low", "high"))
)
```


I computed means for the random sample, the logical vector, the character vector, and the factor vector.
```{r mean_code} 

mean(q1_df$random_samp)
mean(q1_df$logical_vec)
mean(q1_df$character_vec)
mean(q1_df$factor_vec)

## The mean can not be computed for the character, and factoral variables because the values are not numeric/integers. The mean was calculated for the random_samp of uniform distribution and the logical_vec.
```


I attempted to convert the logical, character, and factor vectors to numeric vectors.
```{r eval = FALSE} 

as.numeric(q1_df$logical_vec),
as.numeric(q1_df$character_vec),
as.numeric(q1_df$factor_vec)

## The as.numeric function converts character values to numeric values. In the logical vector, 0 is FALSE and 1 is TRUE. In the factor vector, 2 is Low, 3 is Medium, and 1 is High. Numeric values cannot be assigned to a character variable.
```


I converted character_vec to a factor vector (new_factorvec). Then converted new_factorvec to a numeric vector. I converted factor_vec to a character vector (new_charvec). Then converted new_charvec to a numeric vector.
```{r conversion_code}

new_factorvec = as.factor(q1_df$character_vec)
as.numeric(new_factorvec)
## Since each character variable is unique it attempted to create 10 levels when converting to the factor (categorical) variables. Assigned random numbers to the 10 level factor (categorical) variable.

new_charvec = as.character(q1_df$factor_vec)
as.numeric(new_charvec)

# Converted the factor (categorical) variables to just character variables, as in, did not distinguish the levels. But could not convert from character to nueric because those numeric/integer values cannot be assigned to character variables.

```


Question 2:
I created a data frame (q2_df) comprised of:
x: a random sample of size 1000 from a standard Normal distribution
y: a random sample of size 1000 from a standard Normal distribution
A logical vector (sum_vec) indicating whether the x + y > 0
A numeric vector (num_sumvec) created by coercing the above logical vector
A factor vector (fac_sumvec) created by coercing the above logical vector
```{r question_2_code}
set.seed(168)
q2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  sum_vec = (x + y) > 0,
  num_sumvec = as.numeric(sum_vec),
  fac_sumvec = as.factor(num_sumvec)
)
```

The size of the dataset is `r length(q2_df$fac_sumvec)`. The mean of x is `r mean(q2_df$x)` and the median of x is `r median(q2_df$x)`. The proportion of cases for which the logic vector is "TRUE" is `r mean(q2_df$sum_vec)`



I made a scatterplot of y vs. x using the logical, numeric, and factor variables from the previous part. Then I exported the plot of the logical variable to my project directory.
```{r plots_code}
logic_plot = ggplot(q2_df, aes(x = x, y = y, color = sum_vec)) + geom_point()
num_plot = ggplot(q2_df, aes(x = x, y = y, color = num_sumvec)) + geom_point()
fac_plot = ggplot(q2_df, aes(x = x, y = y, color = fac_sumvec)) + geom_point()

## The color scale of the logical vector and the factor vector are the same, red for FALSE or 0 and blue for TRUE or 1. The color scale of the numeric vector is a gradient that represents the different numeric values with different shades of blue.

ggsave("logic_scatterplot.pdf", plot = logic_plot)

```





