---
title: "P8105_hw1_tj2383"
author: Tanvi Jain
date: 2018-09-15
output: github_document
---

Loaded necessary packages:

```{r}
library(tidyverse)
```

Question 1:
I created a data frame (q1_df) comprised of:

-A random sample (random_samp) of size 10 from a uniform[0, 5] distribution
-A logical vector (logical_vec) indicating whether elements of the sample are greater than 2
-A (length-10) character vector (character_vec)
-A (length-10) factor vector (factor_vec)

```{r question_1_code}
set.seed(168)
q1_df = tibble(
  random_samp = runif(10, min = 0, max = 5),
  logical_vec = random_samp > 2,
  character_vec = c("dog", "cat", "rabbit", "frog", "mouse", "rat", "fly", "hippo", "horse", "bird"),
  factor_vec = factor(c("low", "medium", "high", "medium", "low", "medium", "high", "low", "low", "high"))
)
```

I computed means for the random sample, the logical vector, the character vector, and the factor vector.

```{r mean_code} 
 
mean(random_samp)
mean(logical_vec)
mean(character_vec)
mean(factor_vec)

## The mean can not be computed for the logical, character, and factoral variables because the values are not numeric. The mean was calculated for the random sample of uniform distribution.
```

I attempted to convert the logical, character, and factor vectors to numeric vectors.

```{r eval = FALSE} 
as.numeric(logical_vec),
as.numeric(character_vec),
as.numeric(factor_vec)

## The as.numeric function converts character values to numeric values. In the logical vector, 0 is FALSE and 1 is TRUE. In the factor vector, 2 is Low, 3 is Medium, and 1 is High. Numeric values cannot be assigned to a character variable.
```


I converted character_vec to a factor vector (new_factorvec). Then converted new_factorvec to a numeric vector. I converted factor_vec to a character vector (new_charvec). Then converted new_charvec to a numeric vector.

```{r conversion_code}
new_factorvec = as.factor(character_vec)
as.numeric(new_factorvec)
## Since each character variable is unique it created 10 levels when converting to the factor (categorical) variables. Assigned random numbers to the 10 level factor (categorical) variable.

new_charvec = as.character(factor_vec)
as.numeric(new_charvec)

# Converted the factor (categorical) variables to just character variables, as in, did not distinguish the levels. But could not convert from character to nueric because those numeric values cannot be assigned to character variables.

```


Question 2:
I created a data frame (q2_df) comprised of:
x: a random sample of size 1000 from a standard Normal distribution
y: a random sample of size 1000 from a standard Normal distribution
A logical vector (sum_vec) indicating whether the x + y > 0
A numeric vector (num_sumvec) created by coercing the above logical vector
A factor vector (fac_sumvec) created by coercing the above logical vector

```{r question_2_code}
set.seed(168)
q2_df = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  sum_vec = x + y > 0,
  num_sumvec = as.numeric(sum_vec),
  fac_sumvec = as.factor(num_sumvec)
)
```

The size of the dataset is `r length(fac_sumvec)`. The mean of x is `r mean(x)` and the median of x is `r median(x)`. The proportion of cases for which the logic vector is "TRUE" is `r mean(sum_vec)`


I made a scatterplot of y vs. x using the logical, numeric, and factor variables from the previous part. Then I exported the plot of the logical variable to my project directory.

```{r plots_code}
logic_plot = ggplot(q2_df, aes(x = x, y = y, color = sum_vec)) + geom_point()
num_plot = ggplot(q2_df, aes(x = x, y = y, color = num_sumvec)) + geom_point()
fac_plot = ggplot(q2_df, aes(x = x, y = y, color = fac_sumvec)) + geom_point()

## The color scale of the logical vector and the factor vector are the same, red for FALSE or 0 and blue for TRUE or 1. The color scale of the numeric vector is a gradient that represents the different numeric values with different shades of blue.

ggsave("logic_scatterplot.pdf", plot = logic_plot)

```





